Validation of an Actuator Line Model Coupled to a Dynamic Stall Model for Pitching Motion Characteristic to Vertical Axis Turbine

### 2016-02-29 abstract.tex
Wrap text in introduction

\section{Introduction}
Operating conditions of VAWT are characterized for complex unsteady flows which give a considerable challenge for both to describe using measurements and to represent through simulation tools\cite{huyer1996unsteady}. Moreover, VAWT blades are inherently exposed to cyclic variation in the angle of attack which gives cyclic blade forces and can give material fatigue damage. Accurate modeling of the varying forces is therefore very important for the design of VAWT.\\
The angle of attack in a fixed pitch VAWT is increasing with decreased tip speed ration (TSR), and at low TSR (common during stall regulation), the blades will experience dynamic stall, where the force coefficients for the blade not only depend on the angle of attack, but also on the rate of change of the angle of attack. The aim of this study is to investigate the performance of a Leishman-Beddoes type dynamic stall model when implemented within an actuator line model, for pitching motion typical to a VAWT.

>>>>

\section{Introduction}

Operating conditions of VAWT are characterized for complex unsteady flows which
give a considerable challenge for both to describe using measurements and to
represent through simulation tools\cite{huyer1996unsteady}. Moreover, VAWT
blades are inherently exposed to cyclic variation in the angle of attack which
gives cyclic blade forces and can give material fatigue damage. Accurate
modeling of the varying forces is therefore very important for the design of
VAWT.

The angle of attack in a fixed pitch VAWT is increasing with decreased tip speed
ration (TSR), and at low TSR (common during stall regulation), the blades will
experience dynamic stall, where the force coefficients for the blade not only
depend on the angle of attack, but also on the rate of change of the angle of
attack. The aim of this study is to investigate the performance of a
Leishman-Beddoes type dynamic stall model when implemented within an actuator
line model, for pitching motion typical to a VAWT.
